---
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
  number_sections: true
 
title: ""
author: "[User-764Q](https://github.com/User-764Q)"
date: "`r paste0('Last Run: ', format(Sys.time(), '%A %d-%B-%Y'))`"
params: 
  param1: "Don't Forget about params"

---

<style>

#TOC {
 font-family: Calibri; 
 font-size: 16px;
 border-color: #3D68DF;
 background: #3D68DF;
}

body {
  font-family: Garamond;
  font-size: 16px; 
  border-color: #D0D0D0;
  background-color: #D0D0D0;
  color: #1A1A1A;
}

pre {
  color: #1A1A1A
  background: #D0D0D0;
  background-color: #D0D0D0
  font-family: Calibri; 
  
}

</style>


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(collapse = TRUE)

knitr::opts_chunk$set(warning = FALSE)

knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(include = TRUE)

custom_black <- '1A1A1A'
custom_white <- 'C0C0C0'
custom_grey_dark <- '6F6F6F'
custom_grey_light <- 'B2B2B2'
custom_accent_blue <- '3D6BFF'

```

```{r libraries etc., message = FALSE }

library(gmp)
library(ggthemes)
library(assertive)
library(tidyverse)
library(openxlsx)
library(lubridate)
library(data.table)

```

```{r dataload}



```

# Weekly Data Analysis 1

## Unsupervised Learning LIBS Potatoe Data 

### Introduction

#### Kaggle Potato Laser, Induced, Breakdown Spectroscopy Data

I pulled this data set from [Kaggle]('https://www.kaggle.com/datasets/adolfocobo/libs-potatoes'), it is pretty sparseley described and I have had to make a few assumptions. 

The 'Class Variable' I'm assuming is 'Potatae' so they have lots of measurements of 10 different potatoes, presumably different varieties. 

There are 15,211 measurements, these are measurements of intensity ranging from 'in theory' 0 - 65,535. In practive negative values occur down to about -100 in the first couple of columns. I assume this is just noise but I'll leave it in for now. 

### Reading in the data  

```{r }

column_names <- c('Potatoe', paste0('F-', 1:15211))

raw_data <- read.csv('LIBS.csv', 
                     header = FALSE, col.names = column_names)

# Remove row one because I can't figure out what potatoe it is

cleaned_data <- raw_data %>%
# Remove row one because I can't figure out what potatoe it is
  filter(Potatoe != 'class') %>%
  mutate(Potatoe = as.numeric(Potatoe))

all_rows <- 1:2127 

training_rows <- sample(all_rows, 1000)

other_rows <-  !all_rows %in% training_rows 

samlpe_data <- cleaned_data[other_rows, ] 
  
training_data <- cleaned_data[training_rows, ]

test_data <- cleaned_data[1:100, ]

```

### Principal Component Analysis 

#### Aggregating the data at Potatoe Level

There are so many features in this data set I'm going to use principal component 
analysis PCA to aggregate it up in a sensible way. 

I think I need to do an aggregation of each potatoe or potatoe type before doing the PCA 

I can't figure out an elegant way to do it and I'm worried that this isn't the right method 
as it looses info on the within potate variance for each variable. 

```{r}

potatoes <- cleaned_data %>%
  pull(Potatoe) %>%
  unique()

agg_data <- NULL 

for (potatoe in potatoes) {
  
  this_round <- cleaned_data %>%
    filter(Potatoe == potatoe)
  
  this_row <- colMeans(this_round) 
    
  agg_data <- cbind(agg_data, this_row)
  
}

agg_data <- agg_data %>% t()

```

```{r}

agg_data <- agg_data %>%
  data.frame(row.names = paste0('Tater_', 1:10))

pr_result <- prcomp(agg_data[,2:15212])

pr_result$x <- pr_result$x * -1

pr_result$rotation <- pr_result$rotation * -1

biplot(pr_result, arrow.len = 0)

```

### Hierarchical Clustering 

```{r}

dist_data <- dist(agg_data)

hclust_data <- hclust(dist_data) 

plot(hclust_data)


```

### KMeans Clustering

#### Initial Nieve K-Means Clustering 

Using 10 Clusters 

```{r}
library(factoextra)
library(cluster)

training_data <- training_data %>%
  na.omit()

k_result <- kmeans(training_data[, 2:15212], 10)

fviz_cluster(k_result, data = training_data[, 2:15212])

```

#### Detrimining number of clusters 

```{r}

fviz_nbclust(training_data, kmeans, method = "wss")

```

Lets go with three running the model. 

Three clusters matches the results of H Clustering I did earlier. 

```{r}

k3_train_data <- training_data[, 2:15212] %>%
  na.omit()

k3_result <- kmeans(k3_train_data, 3)

fviz_cluster(k3_result, k3_train_data)

```
#### Seeing how the Potato types map to the clusters

```{r}

mapped_data <- data.frame(cluster = k3_result$cluster, 
                          Potatoe = training_data$Potatoe)

mapped_data <- mapped_data %>% 
  mutate(Potatoe_fct = as.factor(Potatoe))
       
mapped_data$cluster <- k3_result$cluster

mapped_data$Potatoe <- training_data$Potatoe

mapped_data %>%
  ggplot(aes(x = cluster, y = Potatoe, col = Potatoe_fct, fill = Potatoe_fct)) +
  geom_col()

```

This demonstrates a very poor fit between cluster and potatoe. Each of the three clusters seems to have a pretty even spread of the different potatoe types. 

Maybe this has been under fit. 

Lets repeat the above vizualisation with 10 clusters. 

```{r}

k3_train_data <- training_data[, 2:15212] %>%
  na.omit()

k3_result <- kmeans(k3_train_data, 10)

fviz_cluster(k3_result, k3_train_data)

```
#### Seeing how the Potato types map to the clusters

```{r}

mapped_data <- data.frame(cluster = k3_result$cluster, 
                          Potatoe = training_data$Potatoe)

mapped_data <- mapped_data %>% 
  mutate(Potatoe_fct = as.factor(Potatoe)) %>%
  mutate(cluster_fct = as.factor(cluster))
       
mapped_data$cluster <- k3_result$cluster

mapped_data$Potatoe <- training_data$Potatoe 

mapped_data %>%
  ggplot(aes(x = Potatoe_fct, y = cluster_fct, col = cluster_fct, fill = cluster_fct)) +
  geom_col(position = 'stack')

```


```{r}

mapped_data <- data.frame(cluster = k3_result$cluster, 
                          Potatoe = training_data$Potatoe)

mapped_data <- mapped_data %>% 
  mutate(Potatoe_fct = as.factor(Potatoe)) %>%
  mutate(cluster_fct = as.factor(cluster))%>%
  mutate(h_Group = recode(mapped_data$Potatoe, 
                          `5` = '1',
                          `8` = '1',
                          `9` = '1',
                          `10` = '1',
                          `4`= '2', 
                          `7` = '2', 
                          `1` = '2',
                          `2` = '3',
                          `3` = '3',
                          `6` = '3',
                          default = '0'))
       
mapped_data$cluster <- k3_result$cluster

mapped_data$Potatoe <- training_data$Potatoe 

mapped_data %>%
  ggplot(aes(x = Potatoe_fct, y = cluster_fct, col = cluster_fct, fill = cluster_fct)) +
  geom_col(position = 'stack') + 
  facet_wrap(~h_Group)

```

```{r}

k3_train_data <- training_data[, 2:15212] %>%
  na.omit()

k3_result <- kmeans(k3_train_data, 3)

fviz_cluster(k3_result, k3_train_data)

h3_group_data <- training_data %>% 
  mutate(h_Group = recode(Potatoe, 
                          `5` = '1',
                          `8` = '1',
                          `9` = '1',
                          `10` = '1',
                          `4`= '2', 
                          `7` = '2', 
                          `1` = '2',
                          `2` = '3',
                          `3` = '3',
                          `6` = '3',
                          default = '0')) %>%
  select(Potatoe, h_Group)

h3_group_data$cluster <- k3_result$cluster 

h3_group_data <- h3_group_data %>% 
  mutate(cluster = factor(cluster))

ggplot(h3_group_data, aes(x = h_Group, y = cluster, fill = cluster)) +
  geom_col()

```

```{r}

h3_group_data %>% 
  mutate(potatoe_fct = factor(Potatoe, levels = 1:10, order = 1:10)) %>%
ggplot(aes(x = h_Group, 
           y = cluster, 
           fill = potatoe_fct, 
           col = potatoe_fct)) +
  geom_col(position = 'dodge')

```


```{r}

h3_group_data %>% 
  mutate(potatoe_fct = factor(Potatoe, levels = 1:10, order = 1:10)) %>%
ggplot(aes(x = h_Group, 
           y = cluster, 
           fill = potatoe_fct, 
           col = potatoe_fct)) +
  geom_jitter()

```

```{r}

h3_group_data %>% 
  mutate(potatoe_fct = factor(Potatoe, levels = 1:10, order = 1:10)) %>%
ggplot(aes(x = h_Group, 
           y = cluster, 
           fill = cluster, 
           col = cluster)) +
  geom_jitter()

```


```{r}

h3_group_data %>%
  mutate(potatoe_fct = factor(Potatoe, levels = 1:10, order = 1:10)) %>%
  ggplot(aes(x = h_Group, y = cluster, col = h_Group)) + 
  geom_jitter()
  
```

```{r}

cross_tab <- h3_group_data %>%
  group_by(h_Group, cluster) %>%
  summarise(Potatoes = n()) %>%
  pivot_wider(cluster, names_from = h_Group, values_from = Potatoes)

cross_tab

print("Test Success")

```

